Data Prep: IUCN and AquaMaps

### 2016-02-12 data_explore/data_prep.Rmd
got the name crap figured out.  next steps: identify data-poor species etc

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           col_types = 'cdddd', progress = TRUE) %>%
  as.data.frame() %>%
  filter(presence != 5) ### removes extinct subpopulations

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)

@@

spp_am <- read.csv(file.path(dir_data, 'spp_am.csv'), stringsAsFactors = FALSE)

spp_list <- spp_list %>%
  mutate(sm_perc = ifelse(!is.na(sm_perc), 0, sm_perc))
  # filter(!is.na(sm_perc)) %>%
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group') %>%
  left_join(spp_am %>% select(am_sid = speciesid, occurcells),
            by = 'am_sid')

write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'), 
          row.names = FALSE)

>>>>

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE}

spp_am_v_iucn <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
message('Building IUCN species to cell table.  This might take a few minutes.')
iucn_map_files      <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
iucn_spp_cells <- lapply(iucn_map_files, data.table::fread) %>%
  bind_rows()
iucn_spp_cells <- iucn_spp_cells %>%
  rename(iucn_sid = id_no, loiczid = LOICZID) %>%
  mutate(iucn_sid = as.integer(iucn_sid),
         presence = as.integer(presence),
         loiczid  = as.integer(loiczid)) %>%
  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
  filter(presence != 5)
### length(unique(iucn_spp_cells$iucn_sid)) results in 2013 species found...
### some species with no cells reported.  2249 listed in the am v iucn file.

message('Loading AquaMaps species to cell table.  This might take a few minutes.')
am_spp_cells <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
                         showProgress = TRUE)
am_spp_cells <- am_spp_cells %>%
  rename(am_sid = speciesid, prob = probability) %>%
  filter(am_sid %in% spp_am_v_iucn$am_sid)
### length(unique(am_spp_cells$am_sid)) results in 2283 species found - great!

@@

spp_list <- spp_list %>%
  mutate(sm_perc = ifelse(!is.na(sm_perc), 0, sm_perc))
  # filter(!is.na(sm_perc)) %>%
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group') %>%
  left_join(read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
                   stringsAsFactors = FALSE) %>% 
              select(am_sid = speciesid, occurcells),
            by = 'am_sid')

write_csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
